### 2022-04-04 scripts/data-cleaning.Rmd
Initial scripts setup

---
title: "data-cleaning"
output: html_document
date: '2022-04-04'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#### Workspace setup ####
library(knitr)
library(janitor)
library(lubridate)
library(opendatatoronto)
library(tidyverse)
library(tidyr)
library(dplyr)
```

#### Data download ####
```{r}
package <- show_package("e271cdae-8788-4980-96ce-6a5c95bc6618")
package
```
```{r}
package <- list_package_resources("https://open.toronto.ca/dataset/ttc-bus-delay-data/")

get_resource('97b66e92-3587-4436-bd2b-bf12bc0cf683')
```