### 2022-04-04 scripts/data-exploration.Rmd

---
title: "data-exploration"
output: html_document
date: '2022-04-04'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(data.table)
```
```{r}
```