### 2022-04-06 scripts/data-cleaning.Rmd
Data cleaning

---
title: "data-cleaning"
output: html_document
date: '2022-04-04'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(data.table)
library(lubridate)
library(openxlsx)
library(readxl)
library(data.table)
library(janitor)
```
```{r}
# clean 2019 april file specifically - extra field
# april2019 <- data.table(read_excel("../inputs/data/bus/ttc-bus-delay-data-2019.xlsx", 4))
# 
# april2019[, `Incident ID`:= NULL]
# head(april2019)
# write.xlsx(april2019, "../inputs/data/bus/ttc-bus-delay-data-2019-updated.xlsx", overwrite = TRUE, sheetName = 'Apr 2019')
# year2019 <- data.table(read_excel("../inputs/data/bus/ttc-bus-delay-data-2019.xlsx", 4))

# get all files in bus directory
file_list <- list.files("../inputs/data/bus")
# for each file in dir:
#   for each sheet in file:
#     append to df list
#   merge all months in list to single df
# merge years if you want
df_list <- list()

# for each file = for each year
for (i in 1:length(file_list)) {
  filename <- sprintf("../inputs/data/bus/%s", file_list[i])
  sheetnames <- getSheetNames(filename)
  sub_df_list <- list()

  # for each sheet = for each month
  for (j in 1:length(sheetnames)) {
    # if (sheetnames[i] == 'Apr 2019' && file_list[i] == 'ttc-bus-delay-data-2019.xlsx') {
    #   df <- data.table(read_excel(filename, j))
    #   df[, `Incident ID`:= NULL]
    # }
    df <- data.table(read_excel(filename, j))
    # df[, `Report Date` := as.Date(`Report Date`)]
    df[, Time := format(Time, format = "%H:%M:%S")]
    df <- clean_names(df)
    sub_df_list[[j]] <- df
  }

  df_year <- rbindlist(sub_df_list, fill = TRUE)

  df_list[[i]] <- df_year
}
# 
# filename <- "../inputs/data/bus/ttc-bus-delay-data-2014.xlsx"
# sheetnames <- getSheetNames(filename)
# for(i in 1:length(sheetnames)) {
#   print(sheetnames[i])
#   df <- read(excel)
# }
# bus_2014 <- read_excel("../inputs/data/bus/ttc-bus-delay-data-2014.xlsx", 1)
# 
# df <- data.table(bus_2014)
# head(df)
```

#### Checking NAs
```{r}
na_list <- list()

for (i in 1:length(df_list)) {
  na_rows <- sum(!complete.cases(df_list[[i]]))
  all_rows <- count(df_list[[i]])
  na_list[i] <- na_rows / all_rows
}

na_list
```

```{r}
# 15-20% rows have NAs; check their distributions before dropping
df_2014 <- df_list[[1]]

df_2014

df_2014_na <- df_2014[rowSums(is.na(df_2014)) > 0, ]

df_2014_na[rowSums(is.na(df_2014_na)) > 0, ] # 15k rows with a null value

df_2014_na[, vehicle :=NULL]
df_2014_na[, direction :=NULL]

df_2014_na[rowSums(is.na(df_2014_na)) > 0, ] # 158 rows with null value

df_2014_na[complete.cases(df_2014_na), mean(min_delay), by=incident][order(-V1)]
df_2014[complete.cases(df_2014), mean(min_delay), by=incident][order(-V1)]

df_2014[complete.cases(df_2014), mean(min_delay), by=day][order(-V1)]
```
```{r}
for (i in 1:length(df_list)) {
  year_df <- df_list[[i]][, route := as.character(route)]
  p <- year_df[complete.cases(year_df), mean(min_delay), by=route][order(-V1)] %>%
    ggplot(aes(x= reorder(route,-V1), y=V1)) + 
    geom_bar(stat="identity") 
  plot(p)
}

# looks like a few routes are delayed much more often - bad data or reality? explore
```

```{r}
for (i in 1:length(df_list)) {
  df <- df_list[[i]][,.SD, .SDcols = !c('vehicle', 'direction')]
  print(df[complete.cases(df), mean(min_delay), by=day][order(-V1)])
}
```

```{r}
df_list[[1]][, month(`Report Date`)]

df_list[[1]][, mean(`Min Delay`), by=list(Incident, month(`Report Date`))][order(-V1)]

df_list[[1]][, mean(min_delay), by=incident]

mean(df_list[[1]][incident=='Diversion', min_delay][1:10])

df_list[[1]]
```