### 2022-04-06 scripts/data-download.Rmd

---
title: "data-download"
output: html_document
date: '2022-04-04'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#### Workspace setup ####
library(knitr)
library(lubridate)
library(opendatatoronto)
library(tidyverse)
library(tidyr)
library(dplyr)
library(data.table)
library(openxlsx)
```

#### Bus data ####
```{r}
package <- list_package_resources("https://open.toronto.ca/dataset/ttc-bus-delay-data/")

package <- data.table(package)

# ids <- package$id
# package[2, ]
# nrow(package)

for(i in 2:nrow(package)) {
  row <- package[i, ]

  df <- get_resource(row$id)
  write.xlsx(df, sprintf("../inputs/data/bus/%s.xlsx", row$name))
}

# df_2014 <- get_resource('97b66e92-3587-4436-bd2b-bf12bc0cf683')
# write.xlsx(df_2014, "../inputs/data/test-2014.xlsx")
```