### 2022-04-07 scripts/data-exploration.Rmd

---
title: "data-exploration"
output: html_document
date: '2022-04-07'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(data.table)
library(lubridate)
```

# Load data
```{r}
df <- readRDS('../inputs/bus-clean.RDS')

head(df)
```

# Summary stats
## Delay over time
```{r}
df[, month_year := format(as.Date(df$report_date), "%Y-%m")]
df[incident == 'Diversion', 
    mean(min_delay), 
    by=month_year] %>%
    ggplot(aes(x=month_year, y=V1, group=1)) +
    geom_line()
```
## Delay by day of week
```{r}
df[report_date > '2015-01-01', mean(min_delay), by=day] %>%
    ggplot(aes(x= reorder(day,-V1), y=V1)) + 
    geom_bar(stat="identity") 
```
# Quick modelling
```{r}
df2015 <- df[report_date %between% c('2015-01-01', '2015-02-01')]
df2015
model <- lm(min_delay ~ day + incident, data = df2015)
summary(model) # show results
```